---
output: 
  html_document:
    code_folding: show
    highlight: kate
    theme: paper
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(dplyr)
library(knitr)
MRW_clean <- readr::read_csv('assets/Transformed Data/MRW_clean_new_dummies.csv') 
library(broom)
collect.results <- function(models, names){
  
  results <- map2_dfr(.x = models,
                      .y = names,
                      .f = function(x, y){tidy(x) %>% mutate(subset = y)}) %>% 
    as_tibble()
  return(results)
}


plot_new_vs_old <- function(term_type, data, std_error = FALSE ){
  if (std_error == TRUE){
    data <- data %>% 
      mutate(estimate = `std.error`)
  }
  
  new_data <- data %>% 
    filter(term == term_type) %>% 
    filter(type == 'New') %>% 
    mutate(estimate = abs(estimate))
  
  old_data <- data %>% 
    filter(term == term_type) %>% 
    filter(type == 'Original') %>% 
    mutate(estimate =  abs(estimate))
  
  
  p <- plot_ly(new_data,
               x = ~subset,
               y = ~estimate,
               type = 'bar',
               color = ~subset)
  
  q <- plot_ly(old_data,
               x = ~subset,
               y = ~estimate,
               type = 'bar',
               color = ~subset,
               colors = "Set1")
  
  pq <- subplot(p, q, nrows = 1, shareY = TRUE, shareX = TRUE) %>% 
    layout(yaxis = list(title = term_type))
  return(pq)
}
```
In this post we continue our Mankiw, Romer and Weil marathon by visualising countries' saving and depreciation curves, derived from the Solow and Augmented Solow model, using MRW's dataset on economic growth with the help of [Plotly](https://plot.ly/r/) and [Crosstalk](https://rstudio.github.io/crosstalk/).

## Calculating alpha and beta

First, we need to estimate the model to derive empirical estimates of $\alpha$ and $\beta$. MRW calculate both parameters using both a restrict and un-restricted model but choose to stick with the restricted parameter estimates so we do the same.

Running the restricted regressions just involves subtracting our previous regressors from each other. Here, we use `sym` and `!!` to ensure the quasiquotation `filter` expects works (this is analogous to the more lengthy version we used before where we'd code each model separately):
```{r, table_1}
library(purrr)
library(tidyr)

MRW_clean_restricted <- MRW_clean %>% 
  mutate(restricted_regressor = log(s) - log(n_g_d))
restricted_models_old <- c('N',
                           'I',
                           'O') %>% 
  map(sym) %>% 
  map(~filter(MRW_clean_restricted, !!. == 1)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor,
                       data = .))
```


Now, we collect results using `collect.results` defined previously and repeat the above for the Augmented model.
```{r}
restricted_model_results_old <- restricted_models_old %>% 
  collect.results(., names = c("OECD Restricted Classic",
                               "Intermediate Restricted Classic",
                               "Non-Oil Restricted Classic"))

calculate.alpha <- function(estimate){
  alpha <- estimate / (1 + estimate)
  return(alpha)
}

restricted_model_results_old <- restricted_model_results_old %>% 
  mutate(alpha = ifelse((term == 'restricted_regressor'), calculate.alpha(estimate), NA))

## Augmented model
MRW_augmented_restricted <- MRW_clean_restricted %>% 
  mutate(restricted_regressor_hc = log(school / 100) - log(n_g_d))

augmented_restricted_model_old <- c("O",
                                    "I",
                                    "N") %>% 
  map(sym) %>% 
  map(~filter(MRW_augmented_restricted, !!. == 1)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor + restricted_regressor_hc,
          data = .)) %>% 
  collect.results(., names = c("OECD Restricted Augmented",
                               "Intermediate Restricted Augmented",
                               "Non-Oil Restricted Augmented"))

```



Calculating $\alpha$ for the classical Solow model was a pretty simple affair as we know that the regression coefficient is equal to $\frac{\alpha}{1-\alpha}$. The Augmented model is a little trickier since we're solving a set of simultaneous equations. However, the equations describing $\alpha$ and $\beta$ are symmetric so we can just swap round `y` and `x` in the `mutate` function. 

We use `spread` here because we want each estimated coefficient to have its own column - an alternative would be to use some form of `apply` across rows.
```{r}
find.param <- function(y, x){
  param <- (x*y) / (x + x^2 + x*y)
  return(param)
}

augmented_parameters <- augmented_restricted_model_old %>%
  select(term, subset, estimate) %>% 
  filter(term != "(Intercept)") %>% 
  spread(term, estimate) %>% 
  mutate(alpha = find.param(restricted_regressor_hc, restricted_regressor),
         beta = find.param(restricted_regressor, restricted_regressor_hc),
         alpha = round(alpha, 2),
         beta = round(beta, 2)) %>% 
  select(subset, alpha, beta)

classic_parameters <- restricted_model_results_old %>% 
  select(subset, alpha) %>% 
  drop_na() %>% 
  mutate(alpha = round(alpha, 2))

estimated_parameters_old <- bind_rows(classic_parameters,
                                  augmented_parameters) %>% 
  separate(subset, c("subset",
                     "estimation_type",
                     "model_type"),
           sep = " ")
estimated_parameters_old  %>% kable(digits = 2)
```


One of the drawbacks of MRW's data partitioning discussed previously is that countries aren't partitioned into mutually exclusive groups. This is a bit of a pain because it means we effectively have multiple $\alpha$s and $\beta$s for any country which is in more than one group. For instance, the United Kingdom is included in every category by MRW so should we calculate its saving curve with an $\alpha$ of 0.6, 0.59 or 0.36?

To overcome this we can return to the mutually exclusive partitions we created [previously](https://edgongithub.github.io/2018-07-31/Clustering-Mankiw-Romer-Weil-EEG) when we explored clustering the dataset. 

Therefore, we repeat the above but with our new 'club' factor. In hindsight, using more functions would have been a lot quicker here:

```{r}
classic_parameters_new <- c("oil",
                           "developing_no_oil",
                           "intermediate_no_oil",
                           "rich") %>%  
  map(~filter(MRW_clean_restricted, club == .)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor,
                       data = .)) %>% 
  collect.results(., names = c("Oil Restricted Classic",
                               "Developing Restricted Classic",
                               "Intermediate Restricted Classic",
                               "Rich Restricted Classic")) %>% 
  mutate(alpha = ifelse((term == 'restricted_regressor'), calculate.alpha(estimate), NA))

augmented_parameters_new <- c("oil",
                           "developing_no_oil",
                           "intermediate_no_oil",
                           "rich") %>% 
  map(~filter(MRW_augmented_restricted, club == .)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor + restricted_regressor_hc,
                       data = .)) %>% 
  collect.results(., names = c("Oil Restricted Augmented",
                               "Developing Restricted Augmented",
                               "Intermediate Restricted Augmented",
                               "Rich Restricted Augmented")) %>% 
  select(term, subset, estimate) %>% 
  filter(term != "(Intercept)") %>% 
  spread(term, estimate) %>% 
  mutate(alpha = find.param(restricted_regressor_hc, restricted_regressor),
         beta = find.param(restricted_regressor, restricted_regressor_hc),
         alpha = round(alpha, 2),
         beta = round(beta, 2)) %>% 
  select(subset, alpha, beta)

classic_parameters_new <- classic_parameters_new %>% 
  select(subset, alpha) %>% 
  drop_na() %>% 
  mutate(alpha = round(alpha, 2))

estimated_parameters_new <- bind_rows(classic_parameters_new,
                                  augmented_parameters_new) %>% 
  separate(subset, c("subset",
                     "estimation_type",
                     "model_type"),
           sep = " ")
estimated_parameters_new %>% kable(digits = 2)

```

Next, we run a quick sanity check by comparing the estimates. It seems like dropping `Oil` is a good idea as these give pretty crazy results (somewhat predictably).

```{r}
estimated_parameters_new_no_oil <- estimated_parameters_new %>% 
  filter(subset != 'Oil')
comparison_tbl <- bind_rows(estimated_parameters_new_no_oil %>%
                              select(model_type, subset, alpha, beta) %>%
                              mutate(type = 'New'),
                            estimated_parameters_old %>% select(model_type, subset, alpha, beta) %>% 
                              mutate(subset = paste0(subset, ' MRW'),
                                     type = 'Original')) 
comparison_tbl %>%
  head %>% 
  kable(digits = 2)
```
 


We can visualise the differences using Plotly and `plot_new_vs_old` which we also defined in a previous post:

```{r, eval = TRUE}
library(plotly)
comparison_plot_classic <- comparison_tbl %>% 
  filter(model_type == 'Classic') %>% 
  gather(term, estimate, alpha, beta)

plot_alpha_beta <- function(data, beta = FALSE){
  data$subset <- factor(data$subset,
                                   levels = c('Developing',
                                              'Intermediate',
                                              'Rich',
                                              'Non-Oil MRW',
                                              'Intermediate MRW',
                                              'OECD MRW'),
                                   ordered = TRUE)
  
  p <- plot_new_vs_old('alpha', data)
  if (beta == TRUE){
    q <- plot_new_vs_old('beta', data)
    p <- subplot(p, q, nrows = 2, titleY = TRUE)
  }
   return(p)
}

p <- plot_alpha_beta(comparison_plot_classic) %>% 
  layout(showlegend = FALSE,
         title = 'Calculating alpha with new and old partitions, \n Classic Solow')
p
```

<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/23.embed"></iframe> -->

The $\alpha$ estimates are pretty different to MRW's original results which is a little surprising considering the underlying regression estimates seemed to be quite similar when we plotted them in the last post. Furthermore, differentiating the $\alpha$ formula with respect to $x$, our regression estimates, gives us $\frac{1}{(1+x)^2}$ which suggests that $\alpha$ is less sensitive to any given change in $x$ than $x$ itself i.e. $$1 \geq\frac{1}{(1+x)^2} \quad \forall \space  x$$
 
[Edit: This isn't true for $-2 < x < 0$ which invalidates the above somewhat.]
 This could be because the restricted regression set-up we're employing effectively compounds all the estimated coefficient differences into one parameter.
 
 Moving onto the Augmented model and the estimates begin to look similar to Mankiw, Romer and Weil's:

```{r, eval = TRUE}
comparison_plot_augmented <- comparison_tbl %>% 
  filter(model_type == 'Augmented') %>% 
  gather(term, estimate, alpha, beta)

p <- plot_alpha_beta(comparison_plot_augmented, beta = TRUE) %>% 
  layout(showlegend = FALSE,
         title = 'Calculating alpha and beta with new and old partitions, \n Augmented Solow')
p
```

<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/25.embed"></iframe> -->

## Plotting classical saving and depreciation curves 

Now that we have estimated our $\alpha$ and $\beta$ parameters, we can generate the necessary curves.

First, we create a vector of capital levels, $k$, for each country and use `left_join` to include country specific variables such as `s` and `n_g_d`. Next, we use `gsub` to reformat the `club` factor entries so that we can perform another left join and add our estimated $\alpha$s and $\beta$s to the relevant countries depending on which 'club' they're in.

Finally, we calculate the savings curve using the formula $sk^\alpha$ and effective depreciation curve $(n + g + d)k$. As a (limited) test to ensure we've done everything correctly we calculate $k^*$ and $s^*$, the steady state levels of capital accumulation and corresponding savings curve point, which should hopefully correspond to where a country's saving and depreciation curve meet.   

```{r, modelling}
k <- rep(seq(from = 0, to = 20, by = .5), nrow(MRW_clean))
countries <- rep(MRW_clean$country, 20/0.5 + 1)

country_tbl <- tibble('k' = k,
                      'country' = countries) %>% 
  left_join(y = MRW_clean %>% 
              select(`1985`, s, n_g_d, country, club), by = 'country')

estimated_parameters_new$club <- estimated_parameters_new$subset %>% 
  gsub('Oil', 'oil', .) %>% 
  gsub('Developing', 'developing_no_oil', .) %>% 
  gsub('Intermediate', 'intermediate_no_oil', .) %>% 
  gsub('Rich', 'rich', .)

country_tbl <- left_join(country_tbl, estimated_parameters_new %>% 
                           select(club, model_type, alpha, beta),
                         by = 'club')

find.saving.point <- function(k, alpha, s){
  point <- s*k^alpha
  return(point)
}

find.effective.depreciation <- function(k, n = 0, g = 0, d = 0){
  eff_depreciation <- k*(n + g + d)
  return(eff_depreciation)
}

country_tbl <- country_tbl %>% 
  mutate(savings_curve = find.saving.point(k = k,
                                           alpha = alpha,
                                           s = s),
         depreciation_curve = find.effective.depreciation(k = k,
                                                          n = n_g_d),
         k_star = (s / n_g_d) ^ (1 / (1 - alpha)),
         s_star = find.saving.point(k = k_star,
                                    alpha = alpha,
                                    s = s))
country_tbl %>% 
  sample_n(5) %>% 
  kable(digits = 2)
```

We now have all the information we need to plot each countries' curves. To demonstrate the difference between subsets we use `facet_wrap(~club)`. The Crosstalk library lets us highlight individual countries within a plot and compare across multiple plots throughout the post.

First, the classic Solow model savings curves:

~~[Edit: Plotly is having some issues with GitHub Pages' Liquid tags currently so I'm using an iframe workaround that doesn't allow for crosstalk's filter functionality so the plots below are quite messy.]~~

```{r, eval = TRUE}
library(crosstalk)
plot_dat <- country_tbl %>% 
         filter(model_type == "Classic") %>% 
         filter(club != "oil")

plot_dat <- SharedData$new(plot_dat, ~country, group = "Choose a Country:")

p <- ggplot(plot_dat,
            aes(k,
                savings_curve,
                color = country,
                text = country)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~club)
q <- p %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
```

<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/27.embed"></iframe> -->

The differences between the rich, developing and intermediate savings curves demonstrate visually that the Solow model predicts conditional rather than unconditional convergence. The fact that intermediate countries' savings curves appear to be higher than the rich countries, whilst puzzling, is coherent with the Solow model.

Now, we add depreciation curves as well as markers for $k^*$ and $s^*$:
```{r dep_plot, eval = TRUE}
## Now adding depreciation curves and markers
p_dep <- p +
  geom_line(aes(k, depreciation_curve)) +
  geom_point(aes(k_star, s_star))

q <- p_dep %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
```

<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/29.embed"></iframe> -->

The above plot is a little messy, so here we just show the curves' intersection:

```{r, eval = TRUE}
plot_dat <- country_tbl %>% 
         filter(model_type == "Classic") %>% 
         filter(club != "oil") %>% 
         select(country, k_star, s_star, club) %>% 
         distinct()

plot_dat <- SharedData$new(plot_dat, ~country, group = "Choose a Country:")
p <- ggplot(plot_dat,
            aes(k_star, s_star, color = country, text = country)) +
  geom_point() +
  facet_wrap(~club) +
  theme_minimal()
q <- p %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
```
<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/31.embed"></iframe> -->

## Plotting augmented saving curves

Now we can do the same but with the Augmented Solow model: 
```{r augmented plots, echo=FALSE, eval = TRUE}
plot_dat <- country_tbl %>% 
         filter(model_type == "Augmented") %>% 
         filter(club != "oil")

plot_dat <- SharedData$new(plot_dat, ~country, group = "Choose a Country:")

p <- ggplot(plot_dat,
            aes(k,
                savings_curve,
                color = country,
                text = country)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~club)
q <- p %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
```
<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/33.embed"></iframe> -->

Comparing the two together:
```{r, echo=FALSE, eval = TRUE}
plot_dat <- country_tbl %>% 
         filter(club != "oil")

plot_dat <- SharedData$new(plot_dat, ~country, group = "Choose a Country:")

p <- ggplot(plot_dat,
            aes(k,
                savings_curve,
                color = country,
                text = country)) +
  geom_line() +
  theme_minimal() +
  facet_grid(model_type ~ club)
q <- p %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
```
<!-- <iframe width="900" height="800" frameborder="0" scrolling="no" src="//plot.ly/~edjee/35.embed"></iframe> -->
This plot supports MRW's conclusion that human capital can help explain differences in income. The addition of `school` in the regression equation and therefore $\beta$ in the Augmented Solow model now means that rich countries' savings curves appear above the intermediate countries as we'd most likely expect. Furthermore, it seems like the poor countries have gotten poorer.

## Conclusion

We've plotted the savings and depreciation curves of the Solow and Augmented Solow model. Whilst the plots are interactive to a degree, thanks to Plotly and Crosstalk, it'd be interesting to allow users to input their own parameters for $\alpha$, $\beta$, $s$, $(n + g + d)$ - unfortunately this would require Shiny which I'm reluctant to use given GitHub Pages' static hosting.

Again our visualisations and tweaks of MRW's original paper broadly reflect their own findings. Part I of this in depth MRW marathon can be found [here](https://edgongithub.github.io/2018-07-22/Empirics-of-Economic-Growth-Replicated-Revisited-One), followed by [Part II](https://edgongithub.github.io/2018-07-29/Empirics-of-Economic-Growth-Replicated-Revisited-Two) and [Part III.](https://edgongithub.github.io/2018-07-31/Clustering-Mankiw-Romer-Weil-EEG)
