---
output: 
  html_document:
    code_folding: show
    highlight: kate
    theme: paper
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(dplyr)
library(knitr)
MRW_clean <- readr::read_csv('assets/Transformed Data/MRW_clean_new_dummies.csv') 
library(broom)
collect.results <- function(models, names){
  
  results <- map2_dfr(.x = models,
                      .y = names,
                      .f = function(x, y){tidy(x) %>% mutate(subset = y)}) %>% 
    as_tibble()
  return(results)
}


plot_new_vs_old <- function(term_type, data, std_error = FALSE ){
  if (std_error == TRUE){
    data <- data %>% 
      mutate(estimate = `std.error`)
  }
  
  new_data <- data %>% 
    filter(term == term_type) %>% 
    filter(type == 'New') %>% 
    mutate(estimate = abs(estimate))
  
  old_data <- data %>% 
    filter(term == term_type) %>% 
    filter(type == 'Original') %>% 
    mutate(estimate =  abs(estimate))
  
  
  p <- plot_ly(new_data,
               x = ~subset,
               y = ~estimate,
               type = 'bar',
               color = ~subset)
  
  q <- plot_ly(old_data,
               x = ~subset,
               y = ~estimate,
               type = 'bar',
               color = ~subset,
               colors = "Set1")
  
  pq <- subplot(p, q, nrows = 1, shareY = TRUE, shareX = TRUE) %>% 
    layout(yaxis = list(title = term_type))
  return(pq)
}
```
In this post we're going to visualise countries' saving and depreciation curves, derived from the Solow and Augmented Solow model, using Mankiw, Romer and Weil's dataset on economic growth with the help of [Plotly](https://plot.ly/r/) and [Crosstalk](https://rstudio.github.io/crosstalk/).

## Calculating alpha and beta

First, we need to estimate the model to derive empirical estimates of $\alpha$ and $\beta$. MRW calculate both parameters using both a restrict and un-restricted model but choose to stick with the restricted parameter estimates so we do the same.

Running the restricted regressions just involves subtracting our previous regressors from each other. Here, we use `sym` and `!!` to ensure the quasiquotation `filter` expects works (this is analagous to the more lengthy version we used before where we'd code each model separately):
```{r, table_1}
library(purrr)
library(tidyr)

MRW_clean_restricted <- MRW_clean %>% 
  mutate(restricted_regressor = log(s) - log(n_g_d))
restricted_models_old <- c('N',
                           'I',
                           'O') %>% 
  map(sym) %>% 
  map(~filter(MRW_clean_restricted, !!. == 1)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor,
                       data = .))
```


Now, we collect results using `collect.results` defined previously and repeat the above for the Augmented model.
```{r}
restricted_model_results_old <- restricted_models_old %>% 
  collect.results(., names = c("OECD Restricted Classic",
                               "Intermediate Restricted Classic",
                               "Non-Oil Restricted Classic"))

calculate.alpha <- function(estimate){
  alpha <- estimate / (1 + estimate)
  return(alpha)
}

restricted_model_results_old <- restricted_model_results_old %>% 
  mutate(alpha = ifelse((term == 'restricted_regressor'), calculate.alpha(estimate), NA))

## Augmented model
MRW_augmented_restricted <- MRW_clean_restricted %>% 
  mutate(restricted_regressor_hc = log(school / 100) - log(n_g_d))

augmented_restricted_model_old <- c("O",
                                    "I",
                                    "N") %>% 
  map(sym) %>% 
  map(~filter(MRW_augmented_restricted, !!. == 1)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor + restricted_regressor_hc,
          data = .)) %>% 
  collect.results(., names = c("OECD Restricted Augmented",
                               "Intermediate Restricted Augmented",
                               "Non-Oil Restricted Augmented"))

```



Calculating $\alpha$ for the classical Solow model was a pretty simple affair as we know that the regression coefficient is equal to $\frac{\alpha}{1-\alpha}$. The Augmented model is a little trickier since we're solving a set of simultaeneous equations. However, the equations describing $\alpha$ and $\beta$ are symmetric so we can just swap round `y` and `x` in the `mutate` function. 

We use `spread` here because we want each estimated coefficient to have its own column - an alternative would be to use some form of `apply` across rows.
```{r}
find.param <- function(y, x){
  param <- (x*y) / (x + x^2 + x*y)
  return(param)
}

augmented_parameters <- augmented_restricted_model_old %>%
  select(term, subset, estimate) %>% 
  filter(term != "(Intercept)") %>% 
  spread(term, estimate) %>% 
  mutate(alpha = find.param(restricted_regressor_hc, restricted_regressor),
         beta = find.param(restricted_regressor, restricted_regressor_hc),
         alpha = round(alpha, 2),
         beta = round(beta, 2)) %>% 
  select(subset, alpha, beta)

classic_parameters <- restricted_model_results_old %>% 
  select(subset, alpha) %>% 
  drop_na() %>% 
  mutate(alpha = round(alpha, 2))

estimated_parameters_old <- bind_rows(classic_parameters,
                                  augmented_parameters) %>% 
  separate(subset, c("subset",
                     "estimation_type",
                     "model_type"),
           sep = " ")
estimated_parameters_old  %>% kable
```


One of the drawbacks of MRW's data partitioning discussed previously is that countries aren't partitioned into mutually exclusive groups. This is a bit of a pain because it means we effectively have multiple $\alpha$s and $\beta$s for any country which is in more than one group. For instance, the United Kingdom is included in every category by MRW so should we calculate its saving curve with an $\alpha$ of 0.6, 0.59 or 0.36?

To overcome this we can return to the mutually exclusive partitions we created [previously](https://edgongithub.github.io/2018-07-31/Clustering-Mankiw-Romer-Weil-EEG) when we explored clustering the dataset. 

Therefore, we repeat the above but with our new 'club' factor. In hindsight, using more functions would have been a lot quicker here:

```{r}

classic_parameters_new <- c("oil",
                           "developing_no_oil",
                           "intermediate_no_oil",
                           "rich") %>%  
  map(~filter(MRW_clean_restricted, club == .)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor,
                       data = .)) %>% 
  collect.results(., names = c("Oil Restricted Classic",
                               "Developing Restricted Classic",
                               "Intermediate Restricted Classic",
                               "Rich Restricted Classic")) %>% 
  mutate(alpha = ifelse((term == 'restricted_regressor'), calculate.alpha(estimate), NA))

augmented_parameters_new <- c("oil",
                           "developing_no_oil",
                           "intermediate_no_oil",
                           "rich") %>% 
  map(~filter(MRW_augmented_restricted, club == .)) %>% 
  map(~lm(log(`1985`) ~ restricted_regressor + restricted_regressor_hc,
                       data = .)) %>% 
  collect.results(., names = c("Oil Restricted Augmented",
                               "Developing Restricted Augmented",
                               "Intermediate Restricted Augmented",
                               "Rich Restricted Augmented")) %>% 
  select(term, subset, estimate) %>% 
  filter(term != "(Intercept)") %>% 
  spread(term, estimate) %>% 
  mutate(alpha = find.param(restricted_regressor_hc, restricted_regressor),
         beta = find.param(restricted_regressor, restricted_regressor_hc),
         alpha = round(alpha, 2),
         beta = round(beta, 2)) %>% 
  select(subset, alpha, beta)

classic_parameters_new <- classic_parameters_new %>% 
  select(subset, alpha) %>% 
  drop_na() %>% 
  mutate(alpha = round(alpha, 2))

estimated_parameters_new <- bind_rows(classic_parameters_new,
                                  augmented_parameters_new) %>% 
  separate(subset, c("subset",
                     "estimation_type",
                     "model_type"),
           sep = " ")
estimated_parameters_new %>% kable

```

Next, we run a quick sanity check by comparing the estimates. It seems like dropping `Oil` is a good idea as these give pretty crazy results (somewhat predictably). Whilst the estimates aren't identical they are in the same ballpark which is good enough for me:

```{r}
estimated_parameters_new_no_oil <- estimated_parameters_new %>% 
  filter(subset != 'Oil')
comparison_tbl <- bind_rows(estimated_parameters_new_no_oil %>%
                              select(model_type, subset, alpha, beta) %>%
                              mutate(type = 'New'),
                            estimated_parameters_old %>% select(model_type, subset, alpha, beta) %>% 
                              mutate(subset = paste0(subset, ' MRW'),
                                     type = 'Original')) 
comparison_tbl %>%
  head %>% 
  kable
```


Visualising the differences using Plotly and `plot_new_vs_old` which we also defined in a previous post:^[I can't figure out why the bar widths refuse to match but suspect it's something to do with the NA values in the $\beta$ column.]

```{r}
library(plotly)
comparison_plot <- comparison_tbl %>% 
  gather(term, estimate, alpha, beta)

comparison_plot$subset <- factor(comparison_plot$subset,
                                 levels = c('Developing',
                                            'Intermediate',
                                            'Rich',
                                            'Non-Oil MRW',
                                            'Intermediate MRW',
                                            'OECD MRW'),
                                 ordered = TRUE)

p <- plot_new_vs_old('alpha', comparison_plot)
q <- plot_new_vs_old('beta', comparison_plot) 
subplot(p, q, nrows = 2, titleY = TRUE) %>% 
  layout(showlegend = FALSE,
         title = 'Calculating alpha and beta with new and old partitions')
```

## Plotting saving and depreciation curves

Now that we have estimated our $\alpha$ and $\beta$ parameters, we can generate the necessary curves.

First, we create a vector of capital levels, $k$, for each country and use `left_join` to include country specific variables such as `s` and `n_g_d`. Next, we use `gsub` to reformat the `club` factor entries so that we can perform another left join and add our estimated $\alpha$s and $\beta$s to the relevant countries depending on which 'club' they're in.

Finally, we calculate the savings curve using the formula $sk^\alpha$ and effective depreciation curve $(n + g + d)k$. As a (limited) test to ensure we've done everything correctly we calculate $k^*$ and $s^*$, the steady state levels of capital accumulation and corresponding savings curve point, which should hopefully correspond to where a country's saving and depreciation curve meet.   

```{r, modelling}


k <- rep(seq(from = 0, to = 20, by = .5), nrow(MRW_clean))
countries <- rep(MRW_clean$country, 20/0.5 + 1)

country_tbl <- tibble('k' = k,
                      'country' = countries) %>% 
  left_join(y = MRW_clean %>% 
              select(`1985`, s, n_g_d, country, club), by = 'country')


estimated_parameters_new$club <- estimated_parameters_new$subset %>% 
  gsub('Oil', 'oil', .) %>% 
  gsub('Developing', 'developing_no_oil', .) %>% 
  gsub('Intermediate', 'intermediate_no_oil', .) %>% 
  gsub('Rich', 'rich', .)

country_tbl <- left_join(country_tbl, estimated_parameters_new %>% 
                           select(club, model_type, alpha, beta),
                         by = 'club')


find.saving.point <- function(k, alpha, s){
  point <- s*k^alpha
  return(point)
}

find.effective.depreciation <- function(k, n = 0, g = 0, d = 0){
  eff_depreciation <- k*(n + g + d)
  return(eff_depreciation)
}

country_tbl <- country_tbl %>% 
  mutate(savings_curve = find.saving.point(k = k,
                                           alpha = alpha,
                                           s = s),
         depreciation_curve = find.effective.depreciation(k = k,
                                                          n = n_g_d),
         k_star = (s / n_g_d) ^ (1 / (1 - alpha)),
         s_star = find.saving.point(k = k_star,
                                    alpha = alpha,
                                    s = s))
country_tbl %>% 
  sample(5) %>% 
  kable
```

We now have all the information we need to plot each countries' curves. To demonstrate the difference between subsets we use `facet_wrap(~club)`. The crosstalk library lets us highlight individual countries within a plot and compare across multiple plots throughout the post.

First, the classic Solow model savings curves:

```{r}
library(crosstalk)
plot_dat <- country_tbl %>% 
         filter(model_type == "Classic") %>% 
         filter(club != "oil")

plot_dat <- SharedData$new(plot_dat, ~country, group = "Choose a Country:")


p <- ggplot(plot_dat,
            aes(k,
                savings_curve,
                color = country,
                text = country)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~club)
q <- p %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
```


Now, we add depreciation curves as well as markers for $k^*$ and $s^*$:
```{r dep_plot}

## Now adding depreciation curves and markers
p_dep <- p +
  geom_line(aes(k, depreciation_curve)) +
  geom_point(aes(k_star, s_star))

q <- p_dep %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.1) %>% 
  layout(showlegend = FALSE)
q
  

```


The above plot is a little messy, so here we just show the curves' intersection:

```{r}
plot_dat <- country_tbl %>% 
         filter(model_type == "Classic") %>% 
         filter(club != "oil") %>% 
         select(country, k_star, s_star, club) %>% 
         distinct()

plot_dat <- SharedData$new(plot_dat, ~country, group = "Choose a Country:")
p <- ggplot(plot_dat,
            aes(k_star, s_star, color = country, text = country)) +
  geom_point() +
  facet_wrap(~club) +
  theme_minimal()
q <- p %>% 
  ggplotly(tooltip = "text") %>%   
  highlight(on = "plotly_click",
            persistent = TRUE,
            selectize = TRUE,
            defaultValues = c("United Kingdom",
                              "Angola",
                              "Brazil"),
            opacity = 0.3) %>% 
  layout(showlegend = FALSE)
q
```

