---
title: "Goodman-Bacon Decomposition Package in R"
author: "Ed Jee"
date: "24/12/2019"
output:
      html_document:
        code_folding: show
        highlight: kate
        theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center")
```

[Evan](https://twitter.com/flack_evan) and I have released an R package [bacon](https://github.com/evanjflack/bacon) to perform the [Goodman-Bacon decomposition.](https://cdn.vanderbilt.edu/vu-my/wp-content/uploads/sites/2318/2019/07/29170757/ddtiming_7_29_2019.pdf) The paper decomposes diff-in-diff estimates with time-varying treatment into a set of "two-by-twos" comparing treated and untreated groups at each timing stage. Evan did 99\% of the work and it was great fun collaborating on an open source project with him. Below is a very quick example of how the package works.


Installing the package from GitHub (we're not on CRAN just yet) and loading libraries:
```{r}
devtools::install_github("evanjflack/bacon")
library(bacon)
library(ggplot2)
library(dplyr)
```

Perform the decomposition and plot estimates vs weights, using the default `math_reform` dataset included in the package:
```{r}
df_bacon <- bacon(incearn_ln ~ reform_math,
                  data = bacon::math_reform,
                  id_var = "state",
                  time_var = "class")


ggplot(df_bacon) +
  aes(x = weight,
      y = estimate,
      shape = factor(type),
      colour = factor(type)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, 
             linetype = "longdash") +
  labs(x = "Weight",
       y = "Estimate",
       shape = "Type",
       colour = "Type",
       title = "Goodman-Bacon decomposition of diff-in-diff estimates vs weights") +
  theme_minimal() +
  theme(legend.position = "bottom")
```


Plotting estimates from largest to smallest alongside the traditional two-way FE estimate:

```{r}
df_bacon %>% 
  mutate(subgroup = paste0(treated, "_", untreated),
         subgroup = factor(subgroup),
         subgroup = forcats::fct_reorder(subgroup, estimate)) %>% 
  ggplot(aes(x = estimate, 
             y = subgroup,
             size = weight)) +
  geom_point() +
  geom_vline(xintercept = weighted.mean(df_bacon$estimate, df_bacon$weight),
             linetype = "longdash") +
  theme_minimal() +
  labs(size = "Weight",
       y = "Subgroup",
       x = "Estimate",
       title = "Goodman-Bacon diff in diff decomposition",
       subtitle = "Dotted line indicates two-way FE estimate.",
       caption = "Subgroups 99999 correspond to never treated groups")
```


We're always open to feedback and suggestions - just open a [GitHub issue.](https://github.com/evanjflack/bacon/issues)

